---
title: "Casino Site Suitability"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
    social: menu
    theme: flatly
    
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(sf)
options(stringsAsFactors = F) 

#Read in the four shapefiles for the analysis
blocks = st_read('data/Gettysburg_blocks.shp')
gettyBoundary = st_read('data/Gettysburg_boundary.shp')
highways = st_read('data/gettysburgHighways.shp')
nParks = st_read('data/gettysburgNationalParks.shp')

```

### This dashboard is an example of a site suitability analysis done in R. The suitability analysis tutorial was created by [Ken Steif](http://urbanspatialanalysis.com/introduction-to-spatial-site-suitability-analysis-in-r/) and the dashboard was created by [Malle Carrasco-Harris, PhD](https://www.linkedin.com/in/malle-carrasco-harris).

```{r}
library(leaflet)
leaflet() %>%
  addTiles() %>%
  setView(lng = -77.2311, 
          lat = 39.8309, 
          zoom = 12)
```

*** 

In this tutorial, we are considering where a potential new casino should be located in Gettysburg, Pennsylvania. We take into account various decision factors, including:

- the town boundary

- the National Park

- ease of access by thoroughfare

- land zoning

- population density

- land area


### The proposed casino should not be in certain places.

```{r}
#Buffer away from downtown 
gettyBuffer = gettyBoundary %>%
  st_buffer(1609) %>%
  st_difference(., gettyBoundary)

#Buffer away from National Park
npBuffer = nParks %>%
  st_buffer(402)

#Create a map with a boundary around downtown and the national parks
library(mapview)
m = mapview(gettyBoundary,
            col.region = '#EC2810',#red
        alpha.regions = 0.5,
        alpha = 1,
        layer.name = 'Gettysburg Downtown') +
  mapview(gettyBuffer, 
          col.region = '#1C5f15',#green
          layer.name = 'Buffer around Gettysburg Downtown') +
  mapview(npBuffer, 
          col.region = '#EC2810',#red
          layer.name = 'Buffer around National Parks')

m@map %>%
  setView(lng = -77.2311, 
          lat = 39.8309,
          zoom = 12)
```

***
- The proposed casino cannot be downtown. We create a one mile buffer around the town boundary. 

- We also create a quarter mile buffer around the National Park areas.


### The proposed casino should be easy to get to.

```{r}
#Adjacent to highway
hwyBuffer = st_buffer(highways,50)

#Create a map with a boundary around downtown, the national parks, and the highways
m = mapview(gettyBoundary,
            col.region = '#EC2810',#red
        layer.name = 'Gettysburg Downtown') +
  mapview(gettyBuffer, 
          col.region = '#1C5f15',#green
          layer.name = 'Buffer around Gettysburg Downtown') +
  mapview(npBuffer, 
          col.region = '#EC2810',#red
          layer.name = 'Buffer around National Parks')+
  mapview(hwyBuffer, 
          col.region = '#1C5f15',#green
          layer.name = 'Buffer around highways')

m@map %>%
  setView(lng = -77.2311, 
          lat = 39.8309,
          zoom = 12)
```

***
- We consider locations that are easy to get to by creating a 50 meter buffer around the highways. 



### The selected site should meet certain inherent specifications.

```{r}
# Existing population density less than median/Greater than 40 acres/Must be commercial
attributeSelect = blocks %>%
  mutate(area = as.integer(st_area(blocks))* 0.000247105) %>%
  filter(area > 40 &
           pden2000 < median(pden2000) &
           zoning == 2)

#Create a map with the blocks and boundaries. 
m = mapview(gettyBoundary,
            col.region = '#EC2810',#red
            alpha.region = .2,
            layer.name = 'Gettysburg Downtown') +
  mapview(attributeSelect,
          col.region = '#1C5f15',#green
          layer.name = 'Selected Census Blocks') +
  mapview(gettyBuffer,
          col.region = '#1C5f15',#green
          layer.name = 'Buffer around Gettysburg Downtown',
          alpha.regions = .2) +
  mapview(npBuffer,
          col.region = '#EC2810',#red
          layer.name = 'Buffer around National Parks',
          alpha.regions = .2)+
  mapview(hwyBuffer,
          col.region = '#1C5f15',#green
          layer.name = 'Buffer around highways',
          alpha.regions = .2)

m@map %>%
  setView(lng = -77.2311, 
          lat = 39.8309,
          zoom = 12)
```

***
The census block must be:

- zoned for commercial interests

- have lower population density than the Gettysburg median

- be larger than than 40 acres. 


### Selection by location based on decision factors.

```{r}
#Find blocks that are part of the attribute select and inside the gettysburg buffer
#Of those, find those that are not in the national park buffer

dissolve = npBuffer %>%
  group_by(STATE) %>% #There's only one state but this allows us to 'dissolve' away the four polygons into one multipolygon.
  summarize(first(STATE))  #there is no sf package 'dissolve'; use a combination of group_by and summarize

blocksInBuffer = attributeSelect[gettyBuffer,] #Find those dplyr-selected properties that fall within the gettysburg mile buffer. 

#use st_disjoin to find blocks that do not intersect the dissolved Gettysburg Buffer since we DON'T want potential areas in National Parks

noPark = st_disjoint(dissolve, blocksInBuffer, sparse = F)
finalSelection = blocksInBuffer[noPark,] %>% st_transform(4326)


cntr_coords = c(mean(st_coordinates(finalSelection)[,1]),
                mean(st_coordinates(finalSelection)[,2]))
  
leaflet() %>%
  addTiles() %>%
  leafem::addFeatures(finalSelection,
                      weight = 1,
                      fillColor='#df1fca',#gold
                      color = '#A89932',#gold,
                      opacity = 1,
                      fillOpacity = .8) %>%
  setView(lng = -77.2311, 
          lat = 39.8309, 
          zoom = 12) %>%
  addMarkers(lng = cntr_coords[1],
             lat = cntr_coords[2],
          popup = 'Suitable Casino Site')
```

***
We have found blocks that meet the appropriate decision factors in regards to zoning, population, and size. 

We can now limit our search to blocks that are outside of Gettysburg downtown, inside our Gettysburg buffer, outside the National Parks, and close to a highway. 
